---
title: "Analyze CRISPRi growth competition data for growth at high CO2 gas feeds"
author: "Ute Hoffmann (Science For Life Laboratory (KTH), Stockholm, Sweden)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: yes
    number_sections: true
  html_document:
    fig_width: 15
    fig_height: 8
    theme: united
    toc: false
    number_sections: true
---

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
library(ggplot2)
library(ggrepel)
library(ggpubr)
library(tidyverse)
library(magrittr)
library(clusterProfiler)
library(enrichplot)
library(pheatmap)

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Aim of the analysis

Basic visualization of CRISPRi data for cultivation with different levels of CO2 (4% or 30% CO2). Data analysis was performed using nf-core-crispriscreen pipeline (https://github.com/MPUSP/nf-core-crispriscreen).

# Analysis

In a first step, the results given by the Nextflow pipeline are loaded.

```{r load-R-data}
load("../results_controlsgRNAs/fitness/result.Rdata")
num_sgRNAs <- read_tsv("../input/number_sgRNAs_per_target.tsv", col_names = c("sgRNA_target", "num_sgRNAs"))
DESeq_result_table <- left_join(DESeq_result_table, num_sgRNAs)

count_matrix <- read_tsv("../results_controlsgRNAs/prepare/all_counts.tsv")
count_matrix$Gene <- NULL

df_samplesheet <- readr::read_csv("../input/samplesheet_CRISPRi_CO2_Elena.csv", col_types = cols()) %>%
    select(all_of(c("sample", "condition", "replicate", "time", "group", "reference_group"))) %>%
    dplyr::mutate(group = factor(`group`))
df_samplesheet$name <- paste("gen_", df_samplesheet$time, "_r_", df_samplesheet$replicate, sep="")
```

## Diagnostic plot to check if control sgRNAs look ok

Several control sgRNAs are included in the CRISPRi library. These control sgRNAs do not target any specific gene and serve as a control.

```{r diagnostic-plot}
plot_controls_sgRNAs <- DESeq_result_table %>% filter(grepl("ctrl", sgRNA_target)) %>%
  ggplot(aes(x = time, y = log2FoldChange, color = sgRNA_target)) +
  geom_line(linewidth = 1) + geom_point(size = 2) + ylim(-5, 5) + facet_wrap(~ condition, ncol = 4)
print(plot_controls_sgRNAs)
ggsave("../R_results_controlssgRNAs/plot_control_sgRNAs.pdf", plot=plot_controls_sgRNAs, width=12, height=12, units="cm")
```

## Diagnostic plot to check if replicates correlate

This is a diagnostic plot to check if replicates correlate. Samples are clustered according to correlation and the resulting tree is divided into 6 sections according to which samples cluster best. Replicates at different conditions and generations cluster and the 30% samples taken at generations 8 and 10 are most dissimilar to the other samples.

```{r, fig.width = 12, fig.height = 10, warning = FALSE, message=FALSE}
df_counts <- tidyr::pivot_longer(count_matrix,
    cols = 2:ncol(count_matrix),
    names_to = "sample", values_to = "n_reads"
)

# sort
df_counts <- arrange(df_counts, sample)
#df_counts <- left_join(df_samplesheet, df_counts)

df_correlation <- df_counts %>%
    tidyr::pivot_wider(names_from = "sample", values_from = "n_reads") %>%
    dplyr::select(-c(1)) %>%
    cor()

# https://bioinformatics.stackexchange.com/questions/22502/manually-set-range-of-colour-scale-in-pheatmap-in-r
color.divisions <- 100

annotation_days = data.frame(row.names=unique(row.names(df_correlation)), generation=as.character(c(0, 0, 0, 10, 10, 10, 4, 4, 4, 8, 8, 8, 0, 0, 0, 0, 10, 10, 10, 10, 4, 4, 4, 4, 8, 8, 8, 8)), condition=c(rep("30perc",12), rep("4perc", 16)), replicate=as.character(c(rep(c(1,2,3),4), rep(c(4,5,6,7), 4)))) 

# https://stackoverflow.com/questions/41628450/r-pheatmap-change-annotation-colors-and-prevent-graphics-window-from-popping-up
# choose colors for replicate that make difference between the two conditions clearly obvious --> four colours that are similar for HC, three that are similar for LD, let replicate 3 of LD _pop_
# choose gradient of colors for generations
# e.g. Tol from https://davidmathlogic.com/colorblind/#%23D81B60-%231E88E5-%23FFC107-%23004D40
okabe <- c("#f0e442ff", "#e69f00ff", "#d55e00ff", "#cc79a7ff", "#009e73ff", "#56b4e9ff", "#0072b2ff", "#aaaaaaff")
tol <- c("#882255ff","#aa4499ff","#cc6677ff","#ddcc77ff","#88cceeff","#44aa99ff","#117733ff","#332288ff")
okabe_gen <- okabe
names(okabe_gen) <- unique(annotation_days$generation)
okabe_replic <- okabe[1:7]
names(okabe_replic) <- unique(annotation_days$replicate)
annotation_color_list <- list(condition=c("cont"=okabe[7], "LD"=okabe[2]), generation=okabe_gen, replicate=okabe_replic)

okabe <- c("#fff2ecff","#f0e442ff", "#f5a700ff", "#9a4400ff")
okabe_gen <- okabe
names(okabe_gen) <- c("0", "4", "8", "10")
based_on_col <- c("#ff9dd5ff","#cc84aeff","#9a5b83ff","#009affff","#007acbff","#005b98ff","#003d66ff")
names(based_on_col) <- unique(annotation_days$replicate)
annotation_color_list <- list(condition=c("4perc"="#005a96ff", "30perc"="#c57ba5ff"), replicate=based_on_col, generation=okabe_gen)

p <- pheatmap(df_correlation, display_numbers=TRUE, treeheight_col=0, cutree_rows = 6, cutree_cols = 6, annotation_row = annotation_days, annotation_colors = annotation_color_list, breaks = seq(-1,1, length.out=(color.divisions + 1)))
p
ggsave("../R_results_controlssgRNAs/correlation/correlation_samples_clustering.png", plot=p, width=11.5, height=8)
ggsave("../R_results_controlssgRNAs/correlation/correlation_samples_clustering.pdf", plot=p, width=11.5, height=8)

p <- pheatmap(df_correlation, display_numbers=FALSE, treeheight_col=0, cutree_rows = 6, cutree_cols = 6, annotation_row = annotation_days, annotation_colors = annotation_color_list, breaks = seq(-1,1, length.out=(color.divisions + 1)))
p
ggsave("../R_results_controlssgRNAs/correlation/correlation_samples_clustering_woNumbers.png", plot=p, width=11.5, height=8)
ggsave("../R_results_controlssgRNAs/correlation/correlation_samples_clustering_woNumbers.pdf", plot=p, width=11.5, height=8)
```

## Add annotation to results tables

In the following, annotation is added to the results table provided by the Nextflow pipeline. Mapping of the sgRNA targets to slr-locus tags is given in this file, downloaded on 24/02/23: https://github.com/m-jahn/R-notebook-crispri-lib/blob/master/sgRNA_library_V2/data/input/mapping_trivial_names.tsv
The appended annotation is based on Uniprot and Cyanobase, partially edited manually. The table used for annotation was created beginning of 2021. Therefore, it does not include several genes which were only recently characterized.
For a detailed description of all the columns given in the results tables, consult https://mpusp.github.io/nf-core-crispriscreen/output or https://www.biorxiv.org/content/10.1101/2023.02.13.528328v1.full.pdf+htmls

```{r mapping-locusTags}
mapping_gene_locus <- read_tsv("../input/2023-02-24_mapping_trivial_names.tsv", show_col_types=FALSE)
names(mapping_gene_locus) <- c("sgRNA_target", "locus")
DESeq_result_table <- DESeq_result_table %>% left_join(mapping_gene_locus)
```

```{r add-annotation}
annotation <- read_tsv("../input/annotation_locusTags_stand13012021.csv", show_col_types = FALSE)
annotation_2 <- annotation[,c(1,2,3)]
names(annotation_2) <- c("locus", "Gene name","Product")
DESeq_result_table <- DESeq_result_table %>% left_join(annotation_2)
```

```{r save-annotated-files}
write_tsv(DESeq_result_table, file="../R_results_controlssgRNAs/annotated_DESeq_result_table.tsv")
df_reduced_info <- unique(subset(DESeq_result_table, DESeq_result_table$time==8 | DESeq_result_table$time==14)[,c(2,25,26,27,4,20,21,23,24,28)])
write_tsv(df_reduced_info, file="../R_results_controlssgRNAs/Reduced_annotated_DESeq_result_table.tsv")
```

```{r wider-table}
df_red_wide <- pivot_wider(df_reduced_info, names_from=condition, values_from=c(wmean_fitness, sd_fitness, p_fitness_adj, comb_score))
df_red_wide$impact_score <- (df_red_wide$wmean_fitness_CO2_30percent - df_red_wide$wmean_fitness_CO2_4percent)/sqrt(2)
write_tsv(df_red_wide, file="../R_results_controlssgRNAs/Wide_DESeq_result_table.tsv")
```

## Visualization

The weighted mean fitness value combines the values of the different sgRNAs targeting the same gene.
Fitness-fitness plots were created to identify genes which behave differently at different gas conditions. This was performed separately for ncRNAs and protein-coding genes.

### Protein-coding genes

```{r restructure-df}
df_reduced <- unique(subset(DESeq_result_table, DESeq_result_table$time==8)[,c(2,4,20)])
df_red_ncRNAs <- subset(df_reduced, grepl("nc_", df_reduced$sgRNA_target))
df_red_no_ncRNAs <- subset(df_reduced, !grepl("nc_", df_reduced$sgRNA_target))
df_red_wide <- pivot_wider(df_red_no_ncRNAs,names_from="condition", values_from=c("wmean_fitness"))
```

```{r fitness-fitness-plots}
plot_fitness_fitness <- function(df_input, y_axis, y_axis_label, x_axis="CO2_4percent", x_axis_label="Weighted mean fitness 4% CO2", filename_save="../R_results_controlssgRNAs/wfitnes_plot.pdf"){
  df_input$diff <- "NO"
  df_input$diff[(df_input[[x_axis]] - df_input[[y_axis]] > 2.5) | (df_input[[x_axis]] - df_input[[y_axis]] < (-2.5))] <- "YES"
  # prepare labels for plot
  df_input$delabel <- NA
  df_input$delabel[df_input$diff !="NO"] <- df_input$sgRNA_target[df_input$diff != "NO"]
  mycolors <- c("darkblue",  "#d3d3d3b2")
  names(mycolors) <- c("YES", "NO")
  p <- ggplot(data=df_input, aes(x=eval(parse(text=x_axis)), y=eval(parse(text=y_axis)), label=delabel,col=diff)) + geom_point(alpha=0.5, show.legend = FALSE) + 
    theme_light() + labs(y=y_axis_label, x=x_axis_label) + theme(legend.position = "none") + geom_abline(intercept=0,slope=1,linetype="dashed",color="black") + geom_abline(intercept=-2.5, slope=1, linetype="dashed", color="black") + geom_abline(intercept=2.5, slope=1, linetype="dashed", color="black") + scale_colour_manual(values = mycolors) + geom_text_repel(fontface="italic") #+ xlim(-6, +7) +ylim(-6,+7)
  ggsave(filename = filename_save, plot=p, width=12, height=12, units="cm")
return(p)
}

plot_fitness_fitness(df_red_wide, "CO2_30percent", y_axis_label="Weighted mean fitness 30% CO2", filename_save = "../R_results_controlssgRNAs/wfitness.pdf")
```

### ncRNAs

These include antisense RNAs, but also other ncRNAs.

```{r fitness-fitness-plots-ncRNAs}
df_red_wide_ncRNA <- pivot_wider(df_red_ncRNAs,names_from="condition", values_from=c("wmean_fitness"))
plot_fitness_fitness(df_red_wide_ncRNA, "CO2_30percent", y_axis_label="Weighted mean fitness 30% CO2", filename_save = "../R_results_controlssgRNAs/ncRNA_wfitness.pdf")
```

## Growth curves of most interesting targets

```{r plotting-functions-moreColor-CI, echo=FALSE}
lineplot_CVinterval_severalColours_meanlog2 <- function(df_to_plot){
  se_vector <- df_to_plot$sd_log2FoldChange/sqrt(df_to_plot$num_sgRNAs)
  se_vector[is.na(se_vector)] <- df_to_plot$lfcSE[is.na(se_vector)]
  df_to_plot$conf <- qnorm(0.975)*se_vector
  plot_p <- ggplot(df_to_plot, aes(x=time, y=wmean_log2FoldChange, ymin=wmean_log2FoldChange-conf, ymax=wmean_log2FoldChange+conf, group=sgRNA_target, color=factor(sgRNA_target), fill=factor(sgRNA_target))) + geom_line(linewidth=0.5) + geom_ribbon(linetype=0, alpha=0.1) + theme_light() + xlab("Time (generations)") + ylab("Log2FC")
  return(plot_p)
}

lineplot_severalColours_meanlog2 <- function(df_to_plot){
  plot_p <- ggplot(df_to_plot, aes(x=time, y=wmean_log2FoldChange, color=factor(sgRNA_target))) + geom_line(linewidth=0.3) + theme_light() + xlab("Time (generations)") + ylab("Log2FC")
  return(plot_p)
}

lineplot_CVinterval_severalColours_log2 <- function(df_to_plot){
  df_to_plot$conf <- qnorm(0.975)*df_to_plot$lfcSE
  alpha_values <- 1-df_to_plot$weight_lfcSE
  names(alpha_values) <- df_to_plot$sgRNA
  plot_p <- ggplot(df_to_plot, aes(x=time, y=log2FoldChange, ymin=log2FoldChange-conf, ymax=log2FoldChange+conf, group=sgRNA, color=factor(sgRNA), fill=factor(sgRNA))) + new_scale("alpha") + geom_line(linewidth=0.3) + geom_ribbon(linetype=0, aes(alpha=sgRNA)) + scale_alpha_manual(values=alpha_values) + theme_light() + xlab("Time (generations)") + ylab("Log2FC")
  print(unique(alpha_values))
  return(plot_p)
}

lineplot_severalColours_log2 <- function(df_to_plot){
  plot_p <- ggplot(df_to_plot, aes(x=time, y=log2FoldChange, group=sgRNA, color=factor(sgRNA))) + new_scale("alpha") + geom_line(linewidth=0.3) + theme_light() + xlab("Time (generations)") + ylab("Log2FC")
  return(plot_p)
}
```

```{r  message=FALSE}
smaller_subset <- unique(subset(DESeq_result_table, DESeq_result_table$sgRNA_target %in% c("ssl0564", "cpcG", "pmgA", "apqZ", "ctrl1") & DESeq_result_table$condition=="CO2_4percent")[,c("sd_log2FoldChange", "lfcSE", "time", "wmean_log2FoldChange", "sgRNA_target", "sgRNA", "log2FoldChange", "num_sgRNAs")])

colors_vector <- c("apqZ"="#117733ff", "cpcG"="#cc6677ff", "ssl0564"="#ddcc77ff", "ctrl1"="#88cceeff", "pmgA"="#332288ff") # other compatible colors: c("#aa4499ff","#44aa99ff","#117733ff", "#882255ff")
p <- lineplot_CVinterval_severalColours_meanlog2(smaller_subset) + ylim(-6,+4.2) + scale_color_manual(values=colors_vector, na.value="grey40") + scale_fill_manual(values=colors_vector, na.value="grey70")
p
ggsave("../R_results_controlssgRNAs/growthCurves/variants_timeLinePlot_4perc.pdf", plot=p)
ggsave("../R_results_controlssgRNAs/growthCurves/variants_timeLinePlot_4perc.png", plot=p)
```

```{r  message=FALSE}
smaller_subset <- unique(subset(DESeq_result_table, DESeq_result_table$sgRNA_target %in% c("ssl0564", "cpcG", "pmgA", "apqZ", "ctrl1") & DESeq_result_table$condition=="CO2_30percent")[,c("sd_log2FoldChange", "lfcSE", "time", "wmean_log2FoldChange", "sgRNA_target", "sgRNA", "log2FoldChange", "num_sgRNAs")])
p <- lineplot_CVinterval_severalColours_meanlog2(smaller_subset) + ylim(-6,+4.2) + scale_color_manual(values=colors_vector, na.value="grey40") + scale_fill_manual(values=colors_vector, na.value="grey70")
p
ggsave("../R_results_controlssgRNAs/growthCurves/variants_timeLinePlot_30perc.pdf", plot=p)
ggsave("../R_results_controlssgRNAs/growthCurves/variants_timeLinePlot_30perc.png", plot=p)
```

## GSEA

Functional enrichment analyses and gene set enrichment analyses help to check if a certain pathway or specific group of genes is especially affected by a treatment. Here, gene set enrichment analyses were performed for either Gene Ontology terms or KEGG pathways. To perform a gene set enrichment analysis, genes are sorted according to some measure, e.g. the log2FC after a certain time or the calculated fitness. Here, we used the weighted fitness of several sgRNAs as measure.
The mapping of locus tags to Gene Ontology terms was downloaded from UniProt on the 18th Jan. 2024.
There is the possibility to somehow weigh the adjusted p value in these calculation, e.g. by multiplying the weighted mean with the adjusted p value. 
Here, only the first few rows of each table is given. Full tables with all found terms/pathways are available.

```{r GSEA-functional-definitions, echo=FALSE}
### code functional enrichment function
  
# read in term_to_gene and term_to_name for enricher()
term_to_gene <- read.table("../input/2024-01-18_GOterms_slrIdentifier.tsv", sep="\t", header=T)
term_to_gene <- term_to_gene[,c(2,1)]
names(term_to_gene) <- c("GO_ID", "locus_tag")

term_to_name <- read.delim("../input/term_to_name.csv", sep="\t", header=T)
  
columns_to_show  <- c(2,3,4,5,7,8)
columns_to_show_KEGG <- c(1,2,3,4,7)
```

In a first step, GSEAs were calculated for the two different CRISPRi libraries separately. The depletion of essential pathways related to "Ribosomes" or "photosynthesis" is a first good quality measure for a CRISPRi screen.

### 4% CO2

```{r GSEAs-4percent}
DESeq_result_table_4percent <- unique(subset(DESeq_result_table, DESeq_result_table$condition=="CO2_4percent" & DESeq_result_table$time==8)[c("wmean_fitness", "locus")])
geneList <- DESeq_result_table_4percent$wmean_fitness
names(geneList) <- DESeq_result_table_4percent$locus
geneList = sort(geneList, decreasing = TRUE)
set.seed(513)
go_gsea_object <- GSEA(geneList, TERM2GENE = term_to_gene, TERM2NAME=term_to_name, seed=TRUE)
print(head(go_gsea_object)[,columns_to_show])
write.csv(go_gsea_object, "../R_results_controlssgRNAs/GSEA_output/GO_GSEA_CRISPRi_4percent.csv")

set.seed(914)
kegg_gsea_object <- gseKEGG(geneList, organism="syn", minGSSize=10, pvalueCutoff = 0.05, seed=TRUE)
print(head(kegg_gsea_object)[,columns_to_show_KEGG])
write.csv(kegg_gsea_object, "../R_results_controlssgRNAs/GSEA_output/KEGG_GSEA_CRISPRi_4percent.csv")
```

```{r GSEAs-4percent-visualization-etc}
p <- gseaplot2(go_gsea_object, geneSetID =c(3, 10, 15, 18, 21, 22, 23, 24,25,26))
p
ggsave("../R_results_controlssgRNAs/GSEA_output/GO_GSEA_4perc_subset.pdf", plot=p, width=20, height=25, units="cm")
p <- gseaplot2(kegg_gsea_object, geneSetID =1:8)
p
ggsave("../R_results_controlssgRNAs/GSEA_output/KEGG_GSEA_4perc_part1.pdf", plot=p, width=20, height=25, units="cm")
p <- gseaplot2(kegg_gsea_object, geneSetID =9:17)
p
ggsave("../R_results_controlssgRNAs/GSEA_output/KEGG_GSEA_4perc_part2.pdf", plot=p, width=20, height=25, units="cm")
p <- gseaplot2(kegg_gsea_object, geneSetID =18:25)
p
ggsave("../R_results_controlssgRNAs/GSEA_output/KEGG_GSEA_4perc_part3.pdf", plot=p, width=20, height=25, units="cm")
```

### 30% CO2

```{r GSEAs-30percent}
DESeq_result_table_30percent <- unique(subset(DESeq_result_table, DESeq_result_table$condition=="CO2_30percent" & DESeq_result_table$time==8)[c("wmean_fitness", "locus")])
geneList <- DESeq_result_table_30percent$wmean_fitness
names(geneList) <- DESeq_result_table_30percent$locus
geneList = sort(geneList, decreasing = TRUE)
set.seed(513)
go_gsea_object <- GSEA(geneList, TERM2GENE = term_to_gene, TERM2NAME=term_to_name, seed=TRUE)
print(head(go_gsea_object)[,columns_to_show])
write.csv(go_gsea_object, "../R_results_controlssgRNAs/GSEA_output/GO_GSEA_CRISPRi_30percent.csv")

set.seed(914)
kegg_gsea_object <- gseKEGG(geneList, organism="syn", minGSSize=10, pvalueCutoff = 0.05, seed=TRUE)
print(head(kegg_gsea_object)[,columns_to_show_KEGG])
write.csv(kegg_gsea_object, "../R_results_controlssgRNAs/GSEA_output/KEGG_GSEA_CRISPRi_30percent.csv")
```

```{r GSEAs-30percent-visualization-etc}
p <- gseaplot2(go_gsea_object, geneSetID =c(3, 6,9,15,16,17,18,19,23,27,28,29,30,32,33))
p
ggsave("../R_results_controlssgRNAs/GSEA_output/GO_GSEA_30perc_subset.pdf", plot=p, width=20, height=25, units="cm")
p <- gseaplot2(kegg_gsea_object, geneSetID =1:9)
p
ggsave("../R_results_controlssgRNAs/GSEA_output/KEGG_GSEA_30perc_part1.pdf", plot=p, width=20, height=25, units="cm")
p <- gseaplot2(kegg_gsea_object, geneSetID =10:18)
p
ggsave("../R_results_controlssgRNAs/GSEA_output/KEGG_GSEA_30perc_part2.pdf", plot=p, width=20, height=25, units="cm")
p <- gseaplot2(kegg_gsea_object, geneSetID =19:27)
p
ggsave("../R_results_controlssgRNAs/GSEA_output/KEGG_GSEA_30perc_part3.pdf", plot=p, width=20, height=25, units="cm")
p <- gseaplot2(kegg_gsea_object, geneSetID =28:36)
p
ggsave("../R_results_controlssgRNAs/GSEA_output/KEGG_GSEA_30perc_part4.pdf", plot=p, width=20, height=25, units="cm")
p <- gseaplot2(kegg_gsea_object, geneSetID =37:45)
p
ggsave("../R_results_controlssgRNAs/GSEA_output/KEGG_GSEA_30perc_part5.pdf", plot=p, width=20, height=25, units="cm")
```

### Difference between 4% and 30%

In a next step, we tried to check which GO terms or KEGG pathways show a divergent enrichment or depletion in the two libraries. For this, weighted fitness means belonging to the two conditions were subtracted from each other. These differences were used as input for the GSEA.

30% data was subtracted from 4% data. Hence, if a value is high or a pathway enriched, it should have a tendency of being enriched in 4%. The other way round, depletion means a pathway should be more enriched in the 30% data set. These enrichments might not be completely reliable since normalization between both data sets is not perfect.

```{r GSEAs-difference}
df_difference <- unique(subset(DESeq_result_table, DESeq_result_table$time==8 & !is.na(DESeq_result_table$locus))[,c("condition", "wmean_fitness", "locus")])
df_difference_wide <- pivot_wider(df_difference, names_from=condition, values_from=wmean_fitness)
df_difference_wide$difference <- df_difference_wide$CO2_4percent - df_difference_wide$CO2_30percent

df_difference_wide_annotated <- df_difference_wide %>% left_join(annotation_2)
write.csv(df_difference_wide_annotated, "../R_results_controlssgRNAs/fitness_values_differneces_annotated.csv")

geneList <- df_difference_wide$difference
names(geneList) <- df_difference_wide$locus
geneList = sort(geneList, decreasing = TRUE)
set.seed(513)
go_gsea_object <- GSEA(geneList, TERM2GENE = term_to_gene, TERM2NAME=term_to_name, seed=TRUE)
print(head(go_gsea_object)[,columns_to_show])
write.csv(go_gsea_object, "../R_results_controlssgRNAs/GSEA_output/GO_GSEA_difference.csv")

set.seed(914)
kegg_gsea_object <- gseKEGG(geneList, organism="syn", minGSSize=10, pvalueCutoff = 0.05, seed=TRUE)
print(head(kegg_gsea_object)[,columns_to_show_KEGG])
write.csv(kegg_gsea_object, "../R_results_controlssgRNAs/GSEA_output/KEGG_GSEA_difference.csv")
```

Results for the comparison of the two libraries were visualized. These plots are separated in three panels. The lowest panel ("Ranked List Metric") shows the metric according to which the genes were sorted. In this case, this was the weighted fitness mean associated with the different genes. The upper panel shows the running enrichment score of terms/pathways which were enriched/depleted in a statistically significant manner. The middle panel shows where the genes associated with these terms are located within the ranked list of genes in the same color code as used in the upper panel.

```{r GSEAs-difference-visualization-etc}
p <- gseaplot2(go_gsea_object, geneSetID =c(1,3,12,17,18,19))
p
ggsave("../R_results_controlssgRNAs/GSEA_output/GO_GSEA_differences_subset.pdf", plot=p, width=20, height=25, units="cm")
p <- gseaplot2(kegg_gsea_object, geneSetID =1:7)
p
ggsave("../R_results_controlssgRNAs/GSEA_output/KEGG_GSEA_differences_part1.pdf", plot=p, width=20, height=25, units="cm")
p <- gseaplot2(kegg_gsea_object, geneSetID =8:14)
p
ggsave("../R_results_controlssgRNAs/GSEA_output/KEGG_GSEA_differences_part2.pdf", plot=p, width=20, height=25, units="cm")
```

# Session info {.unnumbered}

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
