---
title: "Analyze CRISPRi growth competition data for growth at high CO2 gas feeds, calculate adjusted p values between two conditions"
author: "Ute Hoffmann (Science For Life Laboratory (KTH), Stockholm, Sweden)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: yes
    number_sections: true
  html_document:
    fig_width: 15
    fig_height: 8
    theme: united
    toc: false
    number_sections: true
---

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
library(ggplot2)
library(ggrepel)
library(ggpubr)
library(tidyverse)
library(magrittr)
library(clusterProfiler)
library(enrichplot)
library(pheatmap)

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Aim of the analysis

Calculation of Wilcoxon rank sum test between sgRNA fitness values between 4% CO2 data and 30% CO2 data.

# Analysis

In a first step, the results given by the Nextflow pipeline are loaded.

```{r load-R-data}
load("../results_controlsgRNAs/fitness/result.Rdata")
num_sgRNAs <- read_tsv("../input/number_sgRNAs_per_target.tsv", col_names = c("sgRNA_target", "num_sgRNAs"))
DESeq_result_table <- left_join(DESeq_result_table, num_sgRNAs)

count_matrix <- read_tsv("../results_controlsgRNAs/prepare/all_counts.tsv")
count_matrix$Gene <- NULL

df_samplesheet <- readr::read_csv("../input/samplesheet_CRISPRi_CO2_Elena.csv", col_types = cols()) %>%
    select(all_of(c("sample", "condition", "replicate", "time", "group", "reference_group"))) %>%
    dplyr::mutate(group = factor(`group`))
df_samplesheet$name <- paste("gen_", df_samplesheet$time, "_r_", df_samplesheet$replicate, sep="")
```

```{r wilcox_test}
get_controls <- function(condition, sgRNA_target){
  sgRNA_spec = sgRNA_target
  cond_spec = condition
  filter(DESeq_result_table, condition != cond_spec & sgRNA_target == sgRNA_spec)$fitness
}

DESeq_result_table <- dplyr::left_join(
  DESeq_result_table,
  DESeq_result_table %>%
    dplyr::group_by(sgRNA_target, condition, time) %>%
    dplyr::summarize(
      .groups = "keep",
      # apply Wilcoxon rank sum test against other condition
      p_fitness_condition = stats::wilcox.test(
        x = fitness,
        y = get_controls(condition, sgRNA_target),
        alternative = "two.sided"
        )$p.value,
      # apply Wilcoxon rank sum test against other condition
      p_fitness_condition_2 = stats::wilcox.test(
        x = unique(fitness),
        y = unique(get_controls(condition, sgRNA_target)),
        alternative = "two.sided"
        )$p.value
      ),
  by = c("sgRNA_target", "condition", "time")
  ) %>%
  group_by(condition, time) %>%
  mutate(
    p_fitness_condition_adj = stats::p.adjust(p_fitness_condition, method = "BH"),
    p_fitness_condition_2_adj = stats::p.adjust(p_fitness_condition_2, method = "BH"),
    comb_score = abs(wmean_fitness) * -log10(p_fitness_adj)
    )
```


```{r save}
save(DESeq_result_table, file = "../R_results_controlssgRNAs/result.Rdata")
readr::write_tsv(DESeq_result_table, file = "../R_results_controlssgRNAs/result.tsv")
```


# Add annotation to results tables

In the following, annotation is added to the results table provided by the Nextflow pipeline. Mapping of the sgRNA targets to slr-locus tags is given in this file, downloaded on 24/02/23: https://github.com/m-jahn/R-notebook-crispri-lib/blob/master/sgRNA_library_V2/data/input/mapping_trivial_names.tsv
The appended annotation is based on Uniprot and Cyanobase, partially edited manually. The table used for annotation was created beginning of 2021. Therefore, it does not include several genes which were only recently characterized.
For a detailed description of all the columns given in the results tables, consult https://mpusp.github.io/nf-core-crispriscreen/output or https://www.biorxiv.org/content/10.1101/2023.02.13.528328v1.full.pdf+htmls

```{r mapping-locusTags}
mapping_gene_locus <- read_tsv("../input/2023-02-24_mapping_trivial_names.tsv", show_col_types=FALSE)
names(mapping_gene_locus) <- c("sgRNA_target", "locus")
DESeq_result_table <- DESeq_result_table %>% left_join(mapping_gene_locus)
```

```{r add-annotation}
annotation <- read_tsv("../input/annotation_locusTags_stand13012021.csv", show_col_types = FALSE)
annotation_2 <- annotation[,c(1,2,3)]
names(annotation_2) <- c("locus", "Gene name","Product")
DESeq_result_table <- DESeq_result_table %>% left_join(annotation_2)
```

```{r save-annotated-files}
write_tsv(DESeq_result_table, file="../R_results_controlssgRNAs/annotated_DESeq_result_table_comparisonsConditions.tsv")
df_reduced_info <- unique(subset(DESeq_result_table, DESeq_result_table$time==8 | DESeq_result_table$time==14)[,c(2,25,26,27,4,20,21,23,24,28)])
write_tsv(df_reduced_info, file="../R_results_controlssgRNAs/Reduced_annotated_DESeq_result_table_comparisonsConditions.tsv")
```

```{r wider-table}
df_red_wide <- pivot_wider(df_reduced_info, names_from=condition, values_from=c(wmean_fitness, sd_fitness, p_fitness_adj, comb_score))
df_red_wide$impact_score <- (df_red_wide$wmean_fitness_CO2_30percent - df_red_wide$wmean_fitness_CO2_4percent)/sqrt(2)
write_tsv(df_red_wide, file="../R_results_controlssgRNAs/Wide_DESeq_result_table_comparisonsConditions.tsv")
```

# Session info {.unnumbered}

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
